\documentclass[12pt,a4paper]{letter}
\usepackage[margin=2.5cm]{geometry}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue!60!black}

\signature{David Alfyorov, Igor Shnyukov}
\address{Corresponding author: David Alfyorov\\
\href{mailto:davidich.alfyorov@gmail.com}{davidich.alfyorov@gmail.com}\\
ORCID: 0009-0003-6027-7837}
\date{March 31, 2026}

\begin{document}
\begin{letter}{Editorial Office\\Classical and Quantum Gravity}

\opening{Dear Editors,}

We submit for your consideration the manuscript entitled
\textbf{``Weyl curvature from the Hasse diagram: a parameter-free
bridge formula for causal sets''} by David Alfyorov and Igor
Shnyukov (35~pages, 8~figures).

The paper constructs a discrete observable (CJ) on causal sets that
responds to the electric part of the Weyl tensor and derives an
analytical formula verified against Monte Carlo data with ratio
data/prediction $= 1.016 \pm 0.015$.  To our knowledge, this is
the first discrete causal-set observable shown to be sensitive to
Weyl curvature, extending the Benincasa--Dowker hierarchy from
scalar curvature to the Weyl level.

The main results are:
\begin{itemize}
\item A parameter-free coefficient
  $C_0 = 32\pi^2/(3 \cdot 9! \cdot 45)$ decomposed into five
  factors, each with a distinct geometric origin.
\item Numerical verification on exact pp-wave causal diamonds
  across $N = 500$--$15{,}000$.
\item Seven diagnostic tests including a de~Sitter null test
  (CJ~$= 0$ exactly, 200~seeds) and polarisation independence
  ($M = 50$ paired seeds, $p = 0.43$).
\item Formal verification of all rational coefficient identities
  in Lean~4 (105 sorry-free theorems, three independent proof
  backends).
\item Documentation of five failed approaches and over thirty
  closed spectral routes.
\end{itemize}

The derivation is conditional on two explicitly stated hypotheses
concerning the continuum limit of the stratified estimator.  The
$N^{8/9}$ scaling exponent is consistent with the theoretical
value $8/9$ at the $1\sigma$ level.  All Monte Carlo data,
analysis code, and Lean~4 proof files are publicly available at
\url{https://github.com/davidichalfyorov-wq/sct-theory}
(DOI: 10.5281/zenodo.19135673).

This manuscript has not been published or submitted for
publication elsewhere, and all authors have read and approved the
final version.  The authors declare no conflicts of interest.

We believe the paper is suitable for Classical and Quantum Gravity
given its direct relevance to the causal set programme, to
discrete approaches to quantum gravity, and to the broader
question of how continuum curvature invariants emerge from
discrete causal order.

\closing{Sincerely,}

\end{letter}
\end{document}
