\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,mathtools}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{enumitem}

\definecolor{statusamber}{RGB}{155,96,0}
\definecolor{statusblue}{RGB}{0,72,130}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textsc{SCT Theory --- Prediction Card}}
\fancyhead[R]{\small\textsc{PPN-1}}
\fancyfoot[C]{\thepage}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=green!40!black,
  urlcolor=blue!60!black,
  pdftitle={SCT Prediction Card: Linear Static PPN-1 Sector},
  pdfauthor={David Alfyorov},
}

\title{%
  \textbf{Prediction Card:}\\[4pt]
  \Large Linear Static PPN-1 Sector in SCT Theory
}
\author{David Alfyorov}
\date{March 2026}

\begin{document}
\maketitle


\begin{center}
\small\textit{Credits: Aliaksandr Samatyia contributed research-assistance and workflow support.}
\end{center}

\begin{center}
  \colorbox{statusamber!15}{\parbox{0.78\textwidth}{\centering
    \textbf{\color{statusamber} STATUS: PARTIAL / LINEAR STATIC SECTOR ONLY}\\[2pt]
    \small This card summarizes the part of PPN-1 that is genuinely derived
    from the repaired NT-4a propagator. It does \emph{not} claim a full
    ten-parameter PPN completion.
  }}
\end{center}

\section{Source}

\begin{itemize}[leftmargin=2em]
  \item \textbf{Immediate input:} NT-4a linearized propagator denominators
  \[
    \Pi_{\mathrm{TT}}(z) = 1 + \frac{13}{60} z \hat F_1(z), \qquad
    \Pi_{\mathrm{s}}(z,\xi) = 1 + 6\left(\xi-\frac{1}{6}\right)^2 z \hat F_2(z,\xi).
  \]
  \item \textbf{Phenomenological regime used here:} local Yukawa
  approximation to the exact linear static kernel.
  \item \textbf{Derived quantity:} the radial profile of the linear PPN
  parameter $\gamma(r)=\Psi(r)/\Phi(r)$.
  \item \textbf{Caution:} the exact normalized NT-4a TT denominator develops
  a positive-real-axis zero near $z \approx 2.41484$, so this card should not
  be read as a ghost-freedom statement.
\end{itemize}

\section{Observable Quantity}

The observable targeted in this phase is the weak-field ratio between the
spatial and temporal metric potentials around a static source:
\[
  g_{00} = -1 + 2\Phi(r), \qquad
  g_{ij} = \bigl(1 + 2\Psi(r)\bigr)\delta_{ij}, \qquad
  \gamma(r) = \frac{\Psi(r)}{\Phi(r)}.
\]
This is the quantity most directly constrained by Solar-System light-deflection
and time-delay experiments.

\section{SCT Prediction in the Local Yukawa Approximation}

With
\[
  m_2 = \Lambda\sqrt{\frac{60}{13}}, \qquad
  m_0 = \frac{\Lambda}{\sqrt{6(\xi-1/6)^2}},
\]
the repaired local static potentials are
\[
  \Phi(r) = \frac{GM}{r}\left[1 - \frac{4}{3}e^{-m_2 r} + \frac{1}{3}e^{-m_0 r}\right],
\]
\[
  \Psi(r) = \frac{GM}{r}\left[1 - \frac{2}{3}e^{-m_2 r} - \frac{1}{3}e^{-m_0 r}\right].
\]
Therefore
\begin{equation}
  \boxed{
  \gamma(r)
  =
  \frac{1 - \frac{2}{3}e^{-m_2 r} - \frac{1}{3}e^{-m_0 r}}
       {1 - \frac{4}{3}e^{-m_2 r} + \frac{1}{3}e^{-m_0 r}}
  }.
\end{equation}
At macroscopic distances $r\Lambda \gg 1$, SCT predicts
\[
  \gamma(r) \to 1
  \qquad \text{with Yukawa-suppressed corrections}.
\]

\section{Important Non-Claim}

This phase does \emph{not} predict:
\[
  \beta, \qquad \xi_{\mathrm{PPN}}, \qquad
  \alpha_1,\alpha_2,\alpha_3, \qquad
  \zeta_1,\zeta_2,\zeta_3,\zeta_4.
\]
Those parameters require the nonlinear post-Newtonian field equations and are
explicitly marked \texttt{not\_derived} in the repaired implementation.

\section{Short-Distance Structure}

The local Yukawa approximation has a subtle dependence on $\xi$:
\begin{itemize}[leftmargin=2em]
  \item For $\xi \neq 1/6$, the scalar Yukawa term cancels the leading $1/r$
  singularity in $\Phi(r)$.
  \item At exact conformal coupling $\xi = 1/6$, the scalar channel is absent
  and the local approximation reopens a residual $1/r$ behavior.
\end{itemize}
Accordingly, ``scalar decoupling'' and ``short-distance regularity'' are not
equivalent statements.

\section{Current Experimental Bounds on \texorpdfstring{$\Lambda$}{Lambda}}

\begin{tabular}{lll}
\toprule
Dataset / observable & Condition used & Result \\
\midrule
Cassini time delay & $|\gamma-1| < 2.3\times10^{-5}$ at $1\,\mathrm{AU}$ &
$\Lambda \gtrsim 6.2\times 10^{-18}\,\mathrm{eV}$ \\
Short-range gravity & Yukawa range $< 50\,\mu\mathrm{m}$ &
$\Lambda \gtrsim 1.84\times 10^{-3}\,\mathrm{eV}$ \\
\bottomrule
\end{tabular}

\section{Falsification Logic}

Within the scope of the local static sector, SCT would be in serious tension if:
\begin{enumerate}[label=\arabic*., leftmargin=2em]
  \item a macroscopic non-Yukawa deviation in $\gamma(r)$ were observed where
  the repaired local approximation predicts exponential suppression;
  \item short-range gravity experiments forced the spin-2 Yukawa scale to be
  parametrically larger than any plausible SCT cutoff window;
  \item the future full nonlocal static kernel disagreed qualitatively with
  the local Yukawa approximation already used here.
\end{enumerate}

\section{Status}

\begin{center}
  \colorbox{statusblue!10}{\parbox{0.82\textwidth}{
    \textbf{\color{statusblue} Repaired status (March 2026).}
    The linear static $\gamma(r)$ sector is implemented and tested.
    The full ten-parameter PPN program is \emph{not} complete. This card
    supersedes any earlier wording that claimed $\beta$ or the remaining
    PPN parameters had already been derived.
  }}
\end{center}

\end{document}
