\documentclass[11pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{booktabs}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\HS}{\mathrm{HS}}
\newcommand{\sa}{\mathrm{sa}}
\newcommand{\cF}{\mathcal F}
\newcommand{\cB}{\mathcal B}
\newcommand{\cP}{\mathcal P}
\newcommand{\E}{\mathbb E}
\newcommand{\R}{\mathbb R}
\newcommand{\Tr}{\operatorname{Tr}}
\newcommand{\Law}{\operatorname{Law}}
\newcommand{\Id}{\mathrm{Id}}

\newcommand{\statusproven}{\textcolor{green!50!black}{\textsf{[PROVEN]}}}
\newcommand{\statuscond}{\textcolor{orange!80!black}{\textsf{[CONDITIONAL]}}}
\newcommand{\statusinterp}{\textcolor{blue!60!black}{\textsf{[INTERPRETIVE]}}}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textsc{SCT Theory --- MR-8 Endpoint}}
\fancyhead[R]{\small\thepage}
\renewcommand{\headrulewidth}{0.4pt}

\hypersetup{
  colorlinks=true,
  linkcolor=blue!60!black,
  citecolor=green!40!black,
  urlcolor=blue!60!black,
  pdftitle={MR-8: Nonperturbative Selector Endpoint in Spectral Causal Theory},
  pdfauthor={David Alfyorov},
}

\title{%
  \textbf{MR-8: Nonperturbative Selector Endpoint}\\[4pt]
  \textbf{in Spectral Causal Theory}\\[8pt]
  \large Finite-Rank Tautology, Universal Sufficiency, and the Remaining Loophole
}
\author{David Alfyorov}
\date{April 2026}

\begin{document}
\maketitle

\begin{center}
\end{center}

\begin{abstract}
We record the current mathematical endpoint of MR-8, the problem of a
nonperturbative external selector for Spectral Causal Theory (SCT).  The
main unconditional result is stark: at finite rank, every separating
state-independent selector channel must be injective; every injective
measurable channel with standard Borel codomain is a lossless Borel
re-encoding of the full truncation; and every such injective channel is
automatically sufficient in the sense of the projective criterion of
MR-8 v44.  Therefore the only finite-rank selector towers that can
possibly succeed are the tautological ones equivalent to the identity
truncation itself.  This leaves a single thin loophole: allow
tautological re-encodings as admissible selector channels.  Under the
explicit non-tautology axiom forbidding that loophole, a full no-go
theorem follows immediately.  We separate throughout what is proven, what
is conditional, and what is an archive-grounded interpretation of the
intended MR-8 program.
\end{abstract}

\tableofcontents
\newpage

\section{Scope}

The earlier MR-8 archive established the following background facts.

\begin{enumerate}[leftmargin=2.2em]
  \item The naive spectral-action path integral diverges.
  \item Corrected two-sided Gaussian local completions exist sectorwise.
  \item Global Gaussian gluing fails across generic sectors.
  \item The correct global object is a pro-torsor of local completed
        measure classes together with density-valued observables.
  \item Internal selectors fail.
  \item External selectors on a fixed sector are classified by sector
        weights plus terminal densities.
  \item The post-v44 unresolved problem is the existence of a
        physically motivated finite-rank channel tower, fixed
        \emph{independently of the selected state}, that realizes those
        external states levelwise.
\end{enumerate}

This note concerns only the last point.  It does not revisit the
construction of the torsor/density formalism itself.

\section{Finite-Rank Setup}

Fix one compact-resolvent completed sector, as in v44.  Let
\[
(\Omega,\cF,\mu)
\]
be the ambient completed probability space, and let
\[
\pi_N:\Omega\to\Omega_N
\]
be the level-$N$ truncation map.  Write
\[
\cF_N:=\sigma(\pi_N).
\]

We assume the standard measurable hypotheses already used in v42--v44:

\begin{itemize}[leftmargin=2em]
  \item each $\Omega_N$ is a standard Borel space;
  \item each finite-rank channel codomain $Y_N$ is standard Borel;
  \item ambient channel codomains $Y$ considered below are standard
        Borel;
  \item regular conditional laws exist in the standard way.
\end{itemize}

For the SCT finite-rank truncations relevant here,
\[
\Omega_N \cong \HS_{\sa}(H_N)
\]
for a finite-dimensional Hilbert space $H_N$.  Hence $\Omega_N$ is a
finite-dimensional real vector space and therefore a Polish, in fact
standard Borel, space.

\section{Separating Channels Must Be Injective}

\begin{theorem}[Point-separation obstruction]
\label{thm:separating-injective}
Let $X$ be a standard Borel space and let
\[
B:X\to Y
\]
be any map.  If
\[
\sigma(B)=\cB(X),
\]
then $B$ is injective.
\end{theorem}

\begin{proof}
Assume that $B$ is not injective.  Then there exist $x\neq x'$ in $X$
such that $B(x)=B(x')$.

Because $X$ is standard Borel, the singleton $\{x\}$ is Borel.  If
$\sigma(B)=\cB(X)$, then there exists some set $E\subseteq Y$ such that
\[
\{x\}=B^{-1}(E).
\]
But $B(x)=B(x')$ implies
\[
x\in B^{-1}(E)\iff x'\in B^{-1}(E),
\]
contradicting $x\in\{x\}$ and $x'\notin\{x\}$.  Therefore $B$ must be
injective.
\end{proof}

\begin{corollary}
\label{cor:finite-rank-injective}
If a finite-rank channel
\[
B_N:\Omega_N\to Y_N
\]
is separating,
\[
\sigma(B_N)=\cB(\Omega_N),
\]
then $B_N$ is injective.
\end{corollary}

\begin{proof}
Apply \cref{thm:separating-injective} with $X=\Omega_N$.
\end{proof}

\begin{remark}
This is the exact finite-rank content of the v42 requirement that a
successful selector must act on a measure-determining observable class.
At finite rank, measure determination already forces pointwise
separation.
\end{remark}

\section{Injective Channels Are Tautological Re-Encodings}

We now isolate the only standard descriptive-set-theoretic input used in
this note.

\begin{proposition}[Standard Borel factorization input]
\label{prop:factorization-input}
Let $X,Y,Z$ be standard Borel spaces, let $f:X\to Y$ be measurable, and
let $g:X\to Z$ be $\sigma(f)$-measurable.  Then there exists a
measurable map
\[
h:Y\to Z
\]
such that
\[
g=h\circ f.
\]
\end{proposition}

\begin{remark}
\cref{prop:factorization-input} is the standard measurable factorization
theorem on standard Borel spaces.  We use it as a standard external
input, exactly as one uses disintegration or Lusin--Souslin.
\end{remark}

\begin{theorem}[Finite-rank tautology theorem]
\label{thm:tautology}
Let
\[
B_N:\Omega_N\to Y_N
\]
be measurable, with both $\Omega_N$ and $Y_N$ standard Borel.  If
$B_N$ is injective, then $B_N$ is a Borel re-encoding of the full
truncation: there exists a measurable map
\[
\Psi_N:B_N(\Omega_N)\to \Omega_N
\]
such that
\[
\Psi_N(B_N(x))=x
\qquad
\forall x\in\Omega_N.
\]
\end{theorem}

\begin{proof}
Because $\Omega_N$ is finite-dimensional Polish, choose a Borel
isomorphism
\[
\chi_N:\Omega_N\to \chi_N(\Omega_N)\subseteq \R^{d_N},
\qquad d_N:=\dim_{\R}\Omega_N.
\]
Write
\[
\chi_N=(\chi_{N,1},\dots,\chi_{N,d_N}).
\]
Each coordinate function
\[
\chi_{N,j}:\Omega_N\to \R
\]
is Borel and therefore $\sigma(B_N)$-measurable.  By
\cref{prop:factorization-input}, for each $j$ there exists a measurable
map
\[
f_{N,j}:Y_N\to \R
\]
such that
\[
\chi_{N,j}=f_{N,j}\circ B_N.
\]
Define
\[
F_N:Y_N\to \R^{d_N},
\qquad
F_N:=(f_{N,1},\dots,f_{N,d_N}).
\]
Then
\[
F_N(B_N(x))=\chi_N(x)
\qquad
\forall x\in\Omega_N.
\]
Hence the image $F_N(B_N(\Omega_N))$ lies in $\chi_N(\Omega_N)$, and we
may define
\[
\Psi_N:=\chi_N^{-1}\circ F_N\big|_{B_N(\Omega_N)}.
\]
This map is measurable and satisfies
\[
\Psi_N(B_N(x))
=
\chi_N^{-1}(F_N(B_N(x)))
=
\chi_N^{-1}(\chi_N(x))
=
x.
\]
Therefore $B_N$ is a Borel re-encoding of the full truncation.
\end{proof}

\begin{corollary}
\label{cor:tautology-from-separation}
Every separating finite-rank channel is a tautological re-encoding of the
full truncation.
\end{corollary}

\begin{proof}
Combine \cref{cor:finite-rank-injective} with \cref{thm:tautology}.
\end{proof}

\section{Injective Channels Are Universally Sufficient}

Let
\[
B_N:\Omega_N\to Y_N
\]
be a measurable finite-rank channel, and write
\[
\widetilde B_N:=B_N\circ \pi_N:\Omega\to Y_N.
\]

\begin{theorem}[Injective channels recover the truncation]
\label{thm:recover-truncation}
If $B_N$ is injective, then
\[
\cF_N=\sigma(\widetilde B_N).
\]
\end{theorem}

\begin{proof}
By \cref{thm:tautology}, there exists a measurable inverse
\[
\Psi_N:B_N(\Omega_N)\to\Omega_N
\]
with $\Psi_N(B_N(x))=x$ for all $x\in\Omega_N$.  Therefore
\[
\pi_N(\omega)
=
\Psi_N(B_N(\pi_N(\omega)))
=
\Psi_N(\widetilde B_N(\omega))
\qquad
\forall \omega\in\Omega.
\]
So $\pi_N$ is measurable with respect to $\sigma(\widetilde B_N)$, which
implies
\[
\cF_N=\sigma(\pi_N)\subseteq \sigma(\widetilde B_N).
\]
The reverse inclusion is automatic because
\[
\widetilde B_N=B_N\circ \pi_N
\]
is $\cF_N$-measurable.  Hence the two $\sigma$-fields coincide.
\end{proof}

\begin{theorem}[Universal sufficiency theorem]
\label{thm:universal-sufficiency}
If $B_N$ is injective, then for every ambient channel
\[
B:\Omega\to Y
\]
with standard Borel codomain $Y$, the finite-rank channel $B_N$ is
sufficient for $B$ relative to $\cF_N$.  Equivalently, there exists a
measurable map
\[
K_N:Y_N\to \cP(Y)
\]
such that
\[
\Law_\mu(B\mid \cF_N)
=
K_N\circ \widetilde B_N
\qquad
\mu\text{-a.s.}
\]
\end{theorem}

\begin{proof}
By \cref{thm:recover-truncation},
\[
\cF_N=\sigma(\widetilde B_N).
\]
The regular conditional law
\[
\omega\mapsto \Law_\mu(B\mid \cF_N)(\omega)
\]
is $\cF_N$-measurable and takes values in the standard Borel space
$\cP(Y)$.  Applying \cref{prop:factorization-input} to
$f=\widetilde B_N$ and
$g=\Law_\mu(B\mid \cF_N)$ yields a measurable map
\[
K_N:Y_N\to \cP(Y)
\]
such that
\[
\Law_\mu(B\mid \cF_N)=K_N\circ \widetilde B_N
\qquad
\mu\text{-a.s.}
\]
This is exactly the v44 sufficiency criterion.
\end{proof}

\begin{corollary}
\label{cor:only-tautological-survive}
The only finite-rank selector channels that can possibly satisfy the
v42--v44 burden are the tautological ones equivalent to the identity
truncation.
\end{corollary}

\begin{proof}
If a finite-rank channel can satisfy the v42 burden, it must be
separating; by \cref{cor:tautology-from-separation} it is tautological.
Conversely, by \cref{thm:universal-sufficiency}, every such tautological
channel is automatically sufficient in the v44 sense.
\end{proof}

\section{Consequences for Natural Coarse Channels}

The working-note chain v46--v50 established theorem-level negative
results for the following natural finite-rank selector patterns:

\begin{itemize}[leftmargin=2em]
  \item natural conjugacy-invariant spectral channels;
  \item compression-only boundary channels;
  \item boundary restrictions of equivariant operator functionals;
  \item standard SJ boundary truncations under full basis covariance;
  \item proper-subspace SJ-type truncations under the weaker relative
        phase-covariance hypothesis;
  \item finite families of correlators on a proper test subspace.
\end{itemize}

Together with \cref{cor:only-tautological-survive}, these results show
that all investigated \emph{nontrivial coarse} selector routes fail.  The
only finite-rank survivors are identity-like recodings of the full
truncation itself.

\section{Conditional No-Go Under Non-Tautology}

\begin{definition}[Finite-rank non-tautology axiom]
\label{def:NT}
An admissible finite-rank selector channel
\[
B_N:\Omega_N\to Y_N
\]
is said to satisfy \emph{non-tautology} if it is not Borel-equivalent to
the identity truncation.
\end{definition}

\begin{theorem}[Conditional Option C closure]
\label{thm:conditional-option-c}
Assume that every admissible finite-rank selector channel satisfies
\cref{def:NT}.  Then no external selector tower can satisfy the v42--v44
program.  Equivalently, under non-tautology, Option~C holds.
\end{theorem}

\begin{proof}
Suppose, for contradiction, that an admissible external selector tower
solves the v42--v44 problem.  Then each finite-rank channel must be
separating; otherwise density-gauge ambiguity already survives at finite
rank.  By \cref{cor:tautology-from-separation}, each such channel is
Borel-equivalent to the identity truncation, contradicting
\cref{def:NT}.  Therefore no such admissible tower exists.
\end{proof}

\begin{remark}
The preceding theorem is conditional because non-tautology is an explicit
admissibility axiom rather than a purely measure-theoretic theorem.
\end{remark}

\section{Status of MR-8}

We summarize the endpoint in three layers.

\subsection*{Proven}
\statusproven
\begin{itemize}[leftmargin=2em]
  \item Option~A, the torsor/density endpoint, is mathematically complete.
  \item Every investigated nontrivial coarse finite-rank selector
        pattern fails.
  \item The only finite-rank selector towers that can survive are the
        tautological ones equivalent to the full truncation.
  \item These tautological towers are automatically sufficient in the
        sense of v44.
\end{itemize}

\subsection*{Conditional}
\statuscond
\begin{itemize}[leftmargin=2em]
  \item If tautological finite-rank re-encodings are ruled out by an
        explicit admissibility axiom, then Option~C follows immediately.
\end{itemize}

\subsection*{Interpretive}
\statusinterp
\begin{itemize}[leftmargin=2em]
  \item The language of v44--v45 already treats the target selector as a
        genuine coarse external readout and treats tautological/vacuous
        towers as failures of meaning.
  \item Under that archive-grounded reading, the negative closure is the
        likely intended endpoint of MR-8.
\end{itemize}

\section{Conclusion}

The current mathematical endpoint of MR-8 is a sharp dichotomy.

\begin{quote}
Either one accepts the torsor/density formalism of Option~A, or one
allows only tautological external selector towers that are equivalent to
the full finite-rank truncation itself.
\end{quote}

If tautological re-encodings are excluded as physically vacuous, then the
program closes negatively and Option~C holds.

This is substantially stronger than the original candidate-by-candidate
formulation of the problem.

\end{document}
