\documentclass[11pt,a4paper]{article}

% ============================================================
% SCT Theory - TEST: Internalization and Style Demonstration
% ============================================================

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{enumitem}

\definecolor{sctblue}{RGB}{0,51,153}
\definecolor{verified}{RGB}{0,128,0}
\definecolor{pending}{RGB}{204,102,0}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\small\textsc{SCT Theory --- Derivation Test}}
\fancyhead[R]{\small\textsc{Internalization Check}}
\fancyfoot[C]{\thepage}

\title{%
  \textbf{Style \& Internalization Test:}\\[6pt]
  \Large Spectral Dimension Flow Sketch (NT-3 Prelude)\\[4pt]
  \large Demonstration of Academic Style and Verification Rigor
}
\author{David Alfyorov}
\date{March 2026}

\begin{document}
\maketitle


\begin{center}
\small\textit{Credits: Aliaksandr Samatyia contributed research-assistance and workflow support.}
\end{center}

\begin{center}
  \colorbox{pending!15}{\parbox{0.6\textwidth}{\centering
    \textbf{\color{pending} STATUS: PRELIMINARY TEST}\\[2pt]
    \small Style demonstration only. Not a mathematically verified result.
  }}
\end{center}

\vspace{1em}

% ============================================================
\section{Overview and Purpose}
% ============================================================

This document serves as a compact style and formatting probe for SCT Theory,
showing the expected mathematical structure, reporting conventions, and
verification framing for the upcoming NT-3 phase.

The goal of NT-3 is to compute the flow of the spectral dimension $d_S(\sigma)$
from the IR limit ($d_S \to 4$) to the UV limit ($d_S \to 2$).

% ============================================================
\section{Formal Definition of Spectral Dimension}
% ============================================================

Following Postulate 3 and the logic laid out in \texttt{prediction-04},
the spectral dimension $d_S$ is defined via the return probability $P(\sigma)$
of a diffusion process generated by the Dirac operator $D^2$:

\begin{equation}
  P(\sigma) = \mathrm{Tr}\left( e^{-\sigma D^2} \right)
  \label{eq:heat_trace}
\end{equation}

The dimension itself is the logarithmic derivative:

\begin{equation}
  \boxed{ d_S(\sigma) = -2 \frac{d \ln P(\sigma)}{d \ln \sigma} }
  \label{eq:ds_def}
\end{equation}

% ============================================================
\section{Asymptotic Behavior (Analytical Sketch)}
% ============================================================

\subsection{Infrared Limit ($\sigma \to \infty$)}

In the large-$\sigma$ (low energy) limit, the heat kernel trace is dominated
by the lowest eigenvalues of $D^2$. For a standard Dirac operator on a
commutative 4-dimensional spin manifold, the leading term is:

\begin{equation}
  P(\sigma) \simeq \frac{1}{(4\pi\sigma)^2} \int_M \mathrm{vol} \propto \sigma^{-2}
\end{equation}

Substituting this into Eq.~\eqref{eq:ds_def}:

\begin{equation}
  d_S(\sigma \to \infty) = -2 \frac{d}{d\ln\sigma} (-2 \ln\sigma) = 4
\end{equation}
\textbf{Check:} The IR limit reproduces the macroscopic dimension $d=4$.

\subsection{Ultraviolet Limit ($\sigma \to 0$)}

In the UV, the discrete causal structure of spacetime modifies the
dispersion relation. The effective Dirac operator incorporates higher-order
momentum suppression. Following the Avramidi resummation methods used in NT-1b,
we hypothesize a modified short-distance behavior where $P(\sigma) \propto \sigma^{-1}$.

\begin{equation}
  d_S(\sigma \to 0) = -2 \frac{d}{d\ln\sigma} (-\ln\sigma) = 2
\end{equation}
\textbf{Pending Verification:} This $d_S \to 2$ transition requires rigorous
proof via the full symbolic verification pipeline in the upcoming NT-3 phase.

% ============================================================
\section{Verification Plan}
% ============================================================

To elevate this sketch to a verified \texttt{PASS} status, the following
8-layer pipeline must be executed:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{llc}
  \toprule
  \textbf{Layer} & \textbf{Methodology} & \textbf{Status} \\
  \midrule
  1. Analytic & Evaluate $\lim_{\sigma\to 0}$ and $\lim_{\sigma\to\infty}$ & \textcolor{pending}{PENDING} \\
  2. Numerical & \texttt{mpmath} integration of spectral density & \textcolor{pending}{PENDING} \\
  4. Independent cross-check & Cross-check lattice formulation vs Heat Kernel & \textcolor{pending}{PENDING} \\
  5. Lean 4 & Formal proof of IR limit equivalence & \textcolor{pending}{PENDING} \\
  \bottomrule
\end{tabular}
\end{center}

\vspace{1em}
\noindent
\textit{This document is a formatting and verification-template probe, not a completed derivation.}

\end{document}
